Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-05 appendix.tex
revised

\begin{align}
  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
      r}\right|_0\right]\delta v_x  +
\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{d\delta
    v_z}{dz}\right),\label{gcorr_terms2} \\
\ii \sigma \delta v_x & = \ii k_x W  -
\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
\end{align}
where subscript $0$ denotes evaluation at $r=r_0$.

In the VGSB, the terms proportional to $\hat{g}_c$ are ignored. For a
power-law disk, these radial gradients are $O(r^{-1})$, and they
appear in comparison with terms of $O(k_x)$. The neglected terms 
therefore have a relative magnitude of $O(\epsilon/\khat)$. This is
small for thin disks ($\epsilon\ll1$) and/or small radial wavelengths
($\khat\gg 1$).

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
An important caveat of the VGSB is that it cannot be used in the
adiabatic limit. We explain this by setting  $\beta\to\infty$ in
Eq. \ref{vertiso_gov_nondim}, which can then be written as

@@

However, in the presence of vertical shear $q\neq0$,
Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
$\khat$, which implies the possibility of instability, regardless of
how much $\gamma$ deviates from unity. This is contradicts the second 
Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that
instability in the adiabatic limit requires the disk to be close to
neutrally stratified. %also need real freq^2 in adia limit 
This is because the VGSB retains the vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial disk
structure responsible for it. Nevertheless, we demonstrate below that
this inconsistency is unimportant for the vertical shear instability,
which occurs for $\beta\ll1$.

\subsection{Effect of global radial gradients}

In  Fig. \ref{gcorr_compare}, we demonstrate the effect of the global radial gradient terms
proportional to $\hat{g}_c$ in 
Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} by calculating the
fundamental VSI growth rates using three approaches. We compute growth rates from the dispersion
relation Eq. \ref{relax_disp}, which utilizes the VGSB;  from the
full linear problem (Eq. \ref{lin_mass}---\ref{lin_energy}) with
$\hat{g}_c=0$ (i.e. the linear VGSB equations); and from the full
linear problem with $\hat{g}_c=1$.

All three methods give similar behavior, and growth rates are in close
agreement for $\beta\lesssim 1$. Differences arise for 
$\beta\gtrsim1$, and as $\beta\to\infty$ the VGSB framework gives a
(spurious) positive growth rate as expected from the discussion
above. Inclusion of the global radial gradient terms results in the
correct behavior  ($\nu\to0$ as $\beta\to\infty$). Despite this
caveat, Fig. \ref{gcorr_compare} shows that provided we consider
$\beta\ll1$, then the VGSB framework is an adequate approximation.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
  0cm]{figures/gcorr_compare} 
  \caption{Growth rate of the fundamental VSI mode as a function of
    the thermal relaxation time $\beta$. The `Analytical (VGSB)' curve
    is calculated from the dispersion relation Eq. \ref{relax_disp};
    the `Numerical (VGSB)' curve is obtained by numerically solving
    Eq. \ref{lin_mass}---\ref{lin_energy} with $\hat{g}_c=0$
    (which defines the linear VGSB equations); and the `Numerical
    (Full)' curve is obtained by numerically solving
    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$, which
    accounts for the radial disk structure.  
    The disk parameters are $\Gamma=1.011$,
    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
    wavenumber is $\khat=30$. 
    \label{gcorr_compare}}  
\end{figure}

@@

In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
eliminating variables to obtain a single equation for $\delta v_z$,
Eq. \ref{vertiso_gov}.  This procedure ignores the vertical dependence of
$D=\kappa^2(z) - \sigma^2$. We show here that this has no significant
consequence for thin disks.

In the global disk with $\Gamma=1$ we have
\begin{align}\label{dkappa2}
  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
    r}\frac{\p\Omega^2}{\p z} = -
  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
    \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
\end{align}
For the local problem we may then write
\begin{align}
  \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
\end{align}
where the function $F$ corresponds to the bracket in
Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity.

 Eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
 Eq. \ref{lin_vz} for vertically isothermal disks, retaining the
 vertical dependence of $D$ and including thermal relaxation, we
 obtain 
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
   \left[\left(1-\chi\right) +
     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
   \right\}\delta v_z,
\end{align}
where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
t_c \gamma\right)$. Making the low-frequency approximation with the
replacement $D\to \Omega_k^2$ gives, in terms of dimensionless
variables,
\begin{align}
   &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
    \frac{ \hat{k}^2}{
      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
  \left\{\left(\chi + \ii \epsilon q
      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left[1 - \chi +
      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
\end{align}   
Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
neglecting these terms has no qualitative effect on the discussion in 
\S\ref{analytic_relax}.  This issue does not
arise for isothermal perturbations discussed in \S\ref{iso_discuss}
since in that case we work with a governing equation for $W$ instead.

@@

\subsection{Thin-disk limit}
For thin disks ($\epsilon\ll1$), in which 
$\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
% by expanding the background density and vertical shear profile in powers
% of $z/r$. To lowest order we obtain 
% \begin{align}
%   &\frac{\rho}{\rho_0} \simeq 
%   \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
%   &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
% \end{align} 
%With this approximation and those in \S\ref{integral_relation}, 
Eq. \ref{iso_ode1} becomes  
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \ii q\epsilon
    \hat{k}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
  0.
\end{align}
We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
that derived by \cite{nelson13}, although we have taken a different
route.  
% To complete the problem we must specify boundary conditions. The 
% maximum vertical domain size should be limited by the thin
% disk and low-frequency approximations (i.e. $z\lesssim r$).     
%It is, however, common to take $\mathrm{max}|z|\to\infty$ and 
We impose the kinetic energy remain bounded at infinity. Even though
we have adopted the thin-disk approximation, an infinite
disk is appropriate in the sense that both the true density profile 
and its thin-disk approximation decay rapidly away from the midplane.

@@

\section{Fiducial model for a protoplanetary disk}\label{mmsn}
For results application in \S\ref{application} we use the disk model
described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
has the surface density distribution
\begin{align}\label{mmsn_sigma}
  \Sigma = 2200
  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
\end{align}
and the temperature profile
\begin{align}\label{mmsn_temp}
  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
\end{align}
which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
$\hat{T}$ are dimensionless coefficients used to scale the model
relative to the MMSN. Their nominal values are unity. By assuming a
vertically isothermal disk, we deduce the disk aspect-ratio 
\begin{align}\label{mmsn_epsilon}
  \epsilon =
  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
\end{align}
and the mid-plane density distribution 
\begin{align}
%  \rho_0 = 2.7\times10^{-9}
%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
\rho_0 = 1.7\times10^{-9}
  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
\end{align}
which implies $p=-39/14$. 
%0.033576258
In addition, we
use the opacity model {\bf(reference?)}
 \begin{align}
   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
   \mathrm{cm}^2\,\mathrm{g}
    =
   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
 \end{align}
where the second equality follows from the model temperature profile
above, and $\hat{\kappa}_d$ has similar meaning as $\hat{\Sigma}$ and
$\hat{T}$.

>>>>

\begin{align}
  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
      r}\right|_0\right]\delta v_x  +
\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \hat{g}_c
  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{d\delta
    v_z}{dz}\right),\label{gcorr_terms2} \\
\ii \sigma \delta v_x & = \ii k_x W  -
\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
\end{align}
where subscript $0$ denotes evaluation at $r=r_0$. We recall terms
proportional to $\hat{g}_c$ are associated with the background density
and pressure radial structures, and nominally $\hat{g}_c=1$.

However, in the VGSB these terms are neglected by 
setting $\hat{g}_c=0$. For a power-law disk, these radial gradients
are $O(r^{-1})$, and they appear in comparison with terms of
$O(k_x)$. The neglected terms  therefore have a relative magnitude of
$O(\epsilon/\khat)$. This is small for thin disks ($\epsilon\ll1$)
and/or small radial wavelengths ($\khat\gg 1$).

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
An important caveat of the VGSB is that it cannot be used in the
adiabatic limit. We explain this by considering vertically isothermal
disks and setting  $\beta\to\infty$ in 
Eq. \ref{vertiso_gov_nondim}, which can then be written as

@@

However, in the presence of vertical shear $q\neq0$,
Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
$\khat$, which implies the possibility of instability, regardless of
how much $\gamma$ deviates from unity. This is contradicts the second 
Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that 
instability in a thin, adiabatic disk requires it to be close to 
neutrally stratified, i.e. $|\gamma-1|\ll1$ for vertically isothermal disks
evolved adiabatically. 
%also need real freq^2 in adia limit 
This is because the VGSB retains the vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial
disk structure responsible for it. Nevertheless, we demonstrate below
that this inconsistency is unimportant for the VSI because it occurs
for $\beta\ll1$.

\subsection{Effect of global radial gradients}
In  Fig. \ref{gcorr_compare}, we plot the effect of the global radial gradient terms
proportional to $\hat{g}_c$ in 
Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} by calculating the
fundamental VSI growth rates using three approaches. We compute growth rates from the dispersion
relation Eq. \ref{relax_disp}, which utilizes the VGSB;  from the
full linear problem (Eq. \ref{lin_mass}---\ref{lin_energy}) with
$\hat{g}_c=0$ (i.e. the linear VGSB equations); and from the full
linear problem with $\hat{g}_c=1$.

All three methods give similar behavior, and growth rates are in close
agreement for $\beta\lesssim 1$. Differences arise for 
$\beta\gtrsim1$, and as $\beta\to\infty$ the VGSB framework gives a
(spurious) growth rate as expected from the discussion
above. Inclusion of the global radial gradient terms results in the
correct behavior  ($\nu\to0$ as $\beta\to\infty$). Despite this
caveat, Fig. \ref{gcorr_compare} shows that provided we consider
$\beta\ll1$, then the VGSB framework is an adequate approximation.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
  0cm]{figures/gcorr_compare} 
  \caption{Growth rate of the fundamental VSI mode as a function of
    the thermal relaxation time $\beta$. The `Analytical (VGSB)' curve
    is calculated from the dispersion relation Eq. \ref{relax_disp};
    the `Numerical (VGSB)' curve is obtained by numerically solving
    the linear problem, Eq. \ref{lin_mass}---\ref{lin_energy}, with $\hat{g}_c=0$
    (which defines the linear VGSB equations); and the `Numerical
    (Full)' curve is obtained by numerically solving 
    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$, which
    accounts for the radial disk structure.  
    The disk parameters are $\Gamma=1.011$,
    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
    wavenumber is $\khat=30$. 
    \label{gcorr_compare}}  
\end{figure}

@@

In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
eliminating variables to obtain a single equation for $\delta v_z$ in
vertically isothermal disks, Eq. \ref{vertiso_gov}.  
This procedure ignores the vertical dependence of
$D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
consequence for thin disks.

In the global disk with $\Gamma=1$ we have
\begin{align}\label{dkappa2}
  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
    r}\frac{\p\Omega^2}{\p z} = -
  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
    \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
\end{align}
For the local problem we may then write
\begin{align}
  \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
\end{align}
where the function $F$ corresponds to the bracket in
Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
3+q$ as $|z|\to\infty$. Since we consider $q=O(1)$, $|F|$ is of
order unity.

Now, eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
Eq. \ref{lin_vz} for vertically isothermal disks, retaining the
vertical dependence of $D$ and including thermal relaxation, we
obtain 
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
   \left[\left(1-\chi\right) +
     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
   \right\}\delta v_z,
\end{align}
where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
t_c \gamma\right)$. Making the low-frequency approximation with the
replacement $D\to \Omega_k^2$ gives, in terms of dimensionless
variables,
\begin{align}
   &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
    \frac{ \hat{k}^2}{
      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
  \left\{\left(\chi + \ii \epsilon q
      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left[1 - \chi +
      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
\end{align}   
Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
neglecting these terms has no qualitative effect on the discussion in 
\S\ref{analytic_relax}.  % This issue does not
% arise for isothermal perturbations discussed in \S\ref{iso_discuss}
% since in that case we work with a governing equation for $W$ instead.

@@

\subsection{Explicit solutions in the thin-disk limit}
For thin disks ($\epsilon\ll1$), we write 
$\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
% by expanding the background density and vertical shear profile in powers
% of $z/r$. To lowest order we obtain 
% \begin{align}
%   &\frac{\rho}{\rho_0} \simeq 
%   \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
%   &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
% \end{align} 
%With this approximation and those in \S\ref{integral_relation}, 
so Eq. \ref{iso_ode1} becomes  
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \ii q\epsilon
    \hat{k}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
  0.
\end{align}
We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
that derived by \cite{nelson13}, although we have taken a different
route.  
% To complete the problem we must specify boundary conditions. The 
% maximum vertical domain size should be limited by the thin
% disk and low-frequency approximations (i.e. $z\lesssim r$).     
%It is, however, common to take $\mathrm{max}|z|\to\infty$ and 
We impose the kinetic energy remain bounded at infinity. Even though
we have adopted the thin-disk approximation, an infinite
disk is appropriate in the sense that both the true density profile 
and its thin-disk approximation decay rapidly away from the midplane.

@@

\section{Fiducial model for a protoplanetary disk}\label{mmsn}
For results application in \S\ref{application} we use the disk model
described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
has the surface density distribution
\begin{align}\label{mmsn_sigma}
  \Sigma = 2200
  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
\end{align}
and the temperature profile
\begin{align}\label{mmsn_temp}
  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
\end{align}
which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
$\hat{T}$ are dimensionless coefficients used to scale the model
relative to the MMSN. Their nominal values are unity. By assuming a
vertically isothermal disk, we deduce the disk aspect-ratio 
\begin{align}\label{mmsn_epsilon}
  \epsilon =
  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
\end{align}
and the mid-plane density distribution 
\begin{align}
%  \rho_0 = 2.7\times10^{-9}
%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
\rho_0 = 1.7\times10^{-9}
  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
\end{align}
which implies $p=-39/14$. 
%0.033576258
In addition, we use the opacity model
 \begin{align}
   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
   \mathrm{cm}^2\,\mathrm{g}^{-1}
    =
   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
 \end{align}
\citep{bell94} where the second equality follows from the model
temperature profile above, and $\hat{\kappa}_d$ has similar meaning as
$\hat{\Sigma}$ and $\hat{T}$.
